---
title: "Frequently Asked Questions (FAQs)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Frequently Asked Questions (FAQs)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Why did we build {reportfactory}?
Some projects may require different types of analyses to be run repeatedly over time, for instance due to updates in data and inputs. 

While `rmkardown::render` is becoming a standard for compiling a single analysis document, a number of issues remain: one needs to keep track of different version of the data/inputs, of the analysis code itself, and of different versions of the outputs.

Further, we had a specific requirement as epidemiologists. As we're frequently out in the field with patchy internet connections, we needed something lightweight dependencies-wise so that it could install well in areas where internet connections are unreliable. 

## Why use {reportfactory}?

*reportfactory* aims to make it easy to manage the development, maintainance, and repeated execution of reports. It has functionality to support this by:

- defining a report as an explicitly dated `.Rmd` file
- providing functions to compile all reports, using by default the most recent
  versions
- storing each report output in a separate, time-stamped folder
- maintaining compatibility with basic workflows, i.e. all reports can still be
  directly compiled using `rmarkdown::render()` for testing purposes (although
  one should make sure to remove the outputs afterwards)
- keeping things simple: no configuration files, no handling of potential
  dependencies between reports, no caching
- git-friendly: the factory is compatible with git-based workflows, it keeps core code versioned but saves copies of outputs locally, making it lightweight, but also keeps improvements distributable. 

## What is a "factory"?
Each report "factory" is a structured folder to empower version control in projects with a need for repeated analysis over time. 

## Why not {drake}?
{drake} is a really neat package for reproducible science. However, it has a fairly steep learning curve and most importantly for our use case, it has far more dependencies.

## Can I still use `rmarkdown::render()`?
Note that manual compilation of each document can still be done the usual way,
using `rmarkdown::render` in the source folder; if you do so, make sure you
remove all output files, as they would prevent further updates from the factory.

## How do I refer to external files in reports?
In the `Rmd` reports, all file paths should be referred to using `here()`,
assuming a path from the root directory.

## Where do I put my data and how do I use it in a report?

We recommend storing all data in a `data/` folder in the root directory. When
loading your data in the reports, make sure you use `here`, e.g.:

```{r eval = FALSE}
my_data <- read.csv(here::here("data/linelist_2018-06-11.csv"))
```

## Where do I put scripts for use?
The rationale is the same as for data: store your scripts in a dedicated folder
at the root of the project, e.g. `scripts/`, and source them from R using
`here`, e.g.:

```{r eval = FALSE}
source(here::here("script/big_chunk_of_r_code.R"))
```

## Where do I put functions for use?
{reportfactory} projects can also be packages. Start adding functions in the `R/` directory, ideally by using `usethis::use_r("my_cool_function")` to get you started. You can also document as you go then! Functions created that way can either be sourced like scripts, can be loaded into memory with {devtools}, or installed.

## Where do I put other things to be includedin reports?

Follow the same idea as for data and scripts, as long as you do not alter the
minimum infrastructure (`report_sources/` and other files).
