---
title: "Getting started with reportfactory"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with reportfactory}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r vignettesetup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_knit$set(root.dir = tempdir())
```


reportfactory is a lightweight R package which facilitates a workflow for compiling multiple `.Rmd` reports using `rmarkdown::render()` within a folder. 

## Basic workflow

1. create a new factory using `new_factory()` and move into this new folder
   
2. go to `report_sources/`, write your `.Rmd` report, using the provided
   examples as inspiration; remove the examples files; make sure you use the
   naming conventions explained above, e.g. `foobar_2018-01-25.Rmd`.

3. check your report by compiling the `.Rmd` manually if needed,
   e.g. `rmarkdown::render("foobar_2018-01-25.Rmd")`; once you are happy with the
   results, **make sure you remove all output files from the source folder**
   
4. run `update_reports()` to generate all outputs, or
   `compile_report("foobar_2018-01-25")` if you just want to produce
   time-stamped outputs for this report; check results in the folder
   `report_outputs`

## Example workflow

### Step 1 - Create a new factory
Create and open a new factory. Here, we create the factory with a default template but stay in our current activities (set `move_in` to TRUE to switch projects). Check out more templates at [reconhub/report_factories_templates](https://github.com/reconhub/report_factories_templates).

```{r creation}
library(reportfactory)
factory_proj <- "new_factory"
new_factory(file.path(tempdir(),factory_proj), include_template = TRUE, move_in = FALSE)
```

### Step 2 - Add your reports
Here we've already created some with the `include_template` being set to TRUE (the default). The helper functions below show the state of the factory.

```{r listing}
list_reports(factory_proj)
list_outputs(factory_proj)
list_deps(factory = factory_proj) # list all needed packages
```

### Step 3 - Compile report(s)

#### Step 3a - A single report
The `compile_report()` function can be used to compile a report using its name, or a partial match for its name. This is useful when you're actively working on developing a single report.

```{r compile_one, eval=FALSE}
compile_report("example_report_2019-01-31.Rmd", factory = factory_proj, quiet = TRUE)
```

Use helper functions to see progress.
```{r compile_one_results}
list_outputs(factory_proj)
```

#### Step 3b - Reports with updates
Compile all reports to get their outputs. By default, this just compiles the most recent versions.

```{r compile_all}
update_reports(factory_proj)
list_outputs(factory_proj)
```